---
title: "Take Home Exercise 1"
author: "Ruosong Zhou"
execute: 
  eval: true
  echo: true
  warning: false
---

# Project Brief

The primary goal for this take-home exercise are:

-   to reveal the distribution of Singapore students' performance in mathematics, reading, and science, and

-   to reveal the relationship between these performances with schools, gender and socioeconomic status of the students.

# 1.Data Set

## 1.1 Installing R packages

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse,haven,dplyr) 
```

## 1.2 Data Set

The PISA 2022 database contains the full set of responses from individual students, school principals and parents. In this exercise we are only going to use Student questionnaire data file.

## 1.3 Improting Data

The code chunk below uses read_sas() of haven to import PISA data into R environment

```{r}
#| eval: false
exam_data <- read_sas("data/cy08msp_stu_qqq.sas7bdat")
```

## 1.4 Data Preparation

Since we only interest in the performances of Singapore's students, we are going to filter the CNT to "SGD" only and rewrite it into a new data file to save further loading time.

```{r}
#| eval: false
exam_data_SG <-exam_data %>%
  filter(CNT ==  "SGP")
write_rds(exam_data_SG,
          "data/stu_qqq_SG.rds")
```

Now we read the new data file and keep working on the other parts of data wrangling.

```{r}
stu_qqq_SG <- read_rds("data/stu_qqq_SG.rds")
```

As the data structure shown, we have 6606 observations (Singapore students), and 1279 variables (1279 questions in questionnaire). To compete our task, we certainly do not need all the variables in the data set, we are going to look through it and identify the related variables for our task.

Here's the list of variables we are going to use:

-   CNTSCHID: School ID

-   PVSCIE: Average Performance of science

-   PVREAD: Average Performance of Reading

-   PVMATH: Average Performance of Math

-   ST004D01T: Gender

-   ST258Q01JA: In the past 30 days, how often did a student not eat because there was not enough money to buy food?

-   ST250Q01JA: Do the student have a room of his/her own

-   ST253Q01JA: How many digital devices with screens are a student have

After all the variables were identified we can form out final data set.

```{R}
final_dataset <- stu_qqq_SG %>%
  mutate(
    PV1READ=rowMeans(select(stu_qqq_SG,PV1READ,PV2READ,PV3READ,PV4READ,PV5READ,PV6READ,PV7READ,PV8READ,PV9READ,PV10READ)),
    PV1MATH=rowMeans(select(stu_qqq_SG,PV1MATH,PV2MATH,PV3MATH,PV4MATH,PV5MATH,PV6MATH,PV7MATH,PV8MATH,PV9MATH,PV10MATH)),
    PV1SCIE=rowMeans(select(stu_qqq_SG,PV1SCIE,PV2SCIE,PV3SCIE,PV4SCIE,PV5SCIE,PV6SCIE,PV7SCIE,PV8SCIE,PV9SCIE,PV10SCIE))
    ) %>%

  select(CNTSCHID,PV1READ,PV1SCIE,PV1MATH,ST004D01T,ST258Q01JA,ST250Q01JA,ST253Q01JA)
str(final_dataset)
```

Data quality issue: All the columns are stored as numerical <db1>, but for the columns like school id, using numerical value is not ideal.In order to do the visualization and analysis on the data we have, we have to convert them into character form.

```{r}
final_dataset$CNTSCHID <- as.character(final_dataset$CNTSCHID)
final_dataset$ST250Q01JA <- as.character(final_dataset$ST250Q01JA)
final_dataset$ST004D01T <- as.character(final_dataset$ST004D01T)
```

The last step will be checking missing value and delete them

```{r}
final_dataset <- na.omit(final_dataset)
```

# 2.Data Analysis and Visulization

## 2.1 Performance Distribution

We are going to graph the distribution of performance in each subject to get a clearer view about how Singapore students doing in the exam.

```{r}
#| code-fold: True
#| code-summary: "show code"
p1<- ggplot(data=final_dataset, 
             aes(x = PV1READ)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Reading scores") +
  theme_ipsum()
  
p2<-ggplot(data=final_dataset, 
             aes(x = PV1SCIE)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Science scores") +
  theme_ipsum()
p3<-ggplot(data=final_dataset, 
             aes(x = PV1MATH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Math scores") +
  theme_ipsum()
Perf_dis <- (p1 / p2) | p3
Perf_dis
```

We can notice all the plots have left skewed which indicate the median of the performance bigger than the mean, which means over 50% of students in Singapore can get a grade above the mean performance.

## 2.2 Factors might have an impact on performance

In this section, we are going to examine different factors and their impacts on student performance.

### 2.2.1 School vs. Performance 

In this section we use box plot to illustrate the performance in different school. To see if there is any difference. Since there are too many school in our data set, the visualization on all of them will be difficulty to read, therefore we are going to compare the differences between schools instead.
The code chunk below shows the data preparation phrase
```{r}
average_math_SG <- mean(final_dataset$PV1MATH)
average_scie_SG <- mean(final_dataset$PV1SCIE)
average_read_SG <- mean(final_dataset$PV1READ)

final_dataset_sch <- final_dataset %>%
  group_by(CNTSCHID) %>%
  mutate(average_math_diff = abs(mean(PV1MATH)-average_math_SG),
         average_science_diff = abs(mean(PV1SCIE)-average_scie_SG),
         average_read_diff = abs(mean(PV1READ)-average_read_SG)) %>%
  select(CNTSCHID, average_math_diff, average_science_diff, average_read_diff) %>%
  unique() 
```

```{r}
p4 <-ggplot(data = final_dataset_sch, aes(y = average_math_diff)) +
  geom_boxplot() +
   geom_boxplot(width = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.05) +
  stat_summary(aes(x = 0),
               geom = "point",       
               fun = "mean",         
               colour = "red", 
               size = 1) +
  scale_y_continuous(NULL, limits = c(0,220)) +
  scale_x_continuous("average math diff", breaks = NULL) +
  theme_minimal()
    
p5 <-ggplot(data = final_dataset_sch, aes(y = average_science_diff)) +
  geom_boxplot() +
   geom_boxplot(width = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.05) +
  stat_summary(aes(x = 0),
               geom = "point",       
               fun = "mean",         
               colour = "red", 
               size = 1) +
  scale_y_continuous(NULL, limits = c(0,220)) +
  scale_x_continuous("average science diff", breaks = NULL) +
  theme_minimal()
p6<-ggplot(data = final_dataset_sch, aes(y = average_read_diff)) +
  geom_boxplot() +
   geom_boxplot(width = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.05) +
  stat_summary(aes(x = 0),
               geom = "point",       
               fun = "mean",         
               colour = "red", 
               size = 1) +
  scale_y_continuous(NULL, limits = c(0,220)) +
  scale_x_continuous("average read diff", breaks = NULL) +
  theme_minimal()
school_impact<-p4|p5|p6
school_impact
```

In the graph, we can conclude that there is huge difference on performance across different school.The difference between school can range up to 220. Therefore, we can conclude that choosing different school could definitely affect students' performance.

### 2.2.2 Gender vs. Performance

Still, we are going to apply box plot to see how girls and boys doing in their performance.

```{r}
p7<-ggplot(data = final_dataset, aes(x =ST004D01T , y = PV1READ)) +
  geom_boxplot() +
  ggtitle("Impact of Gender on Read Performance")+
  theme(plot.title = element_text(size = 6))
p8<-ggplot(data = final_dataset, aes(x =ST004D01T , y = PV1SCIE)) +
  geom_boxplot() +
  ggtitle("Impact of Gender on Science Performance")+
  theme(plot.title = element_text(size = 6))
p9<-ggplot(data = final_dataset, aes(x =ST004D01T , y = PV1MATH)) +
  geom_boxplot() +
  ggtitle("Impact of Gender on Math Performance")+
  theme(plot.title = element_text(size = 6))
gender_impact<-p7+p8+p9
gender_impact
```

As shown in the graph, no serious difference detected. Therefore, we can conclude that the gender will not be a factor that cause the differences in performance.

### 2.2.3 Having Own room or not vs. Performance

As the children grown, they will start to require for the "private space", therefore we are going to check if having private space can actually cause impact on student performance or not.

```{r}
p10<-ggplot(data = final_dataset, aes(x =ST250Q01JA , y = PV1READ)) +
  geom_boxplot() +
  ggtitle("Private Space vs. Read Performance")+
  theme(plot.title = element_text(size = 6))
p11<-ggplot(data = final_dataset, aes(x =ST250Q01JA , y = PV1SCIE)) +
  geom_boxplot() +
  ggtitle("Private Space vs. Science Performance")+
  theme(plot.title = element_text(size = 6))
p12<-ggplot(data = final_dataset, aes(x =ST250Q01JA , y = PV1MATH)) +
  geom_boxplot() +
  ggtitle("Private Space vs. Math Performance")+
  theme(plot.title = element_text(size = 6))
pri_impact<-p10+p11+p12
pri_impact
```

As the box plot shown, the mean performance of students who having their own room are slightly higher than those students who do not. Therefore, we can conclude that having own room can be a factor that cause the difference in student's performance.

### 2.2.4 Starving vs. Performance

Question code ST258Q01JA is the a make me shock the most. I do not know where this questionnaire is taking place but certainly a developed region. But still, people still get staved here. The question is "In the past 30 days, how often did a student not eat because there was not enough money to buy food?". I think this poverty will definitely affect the performance in a good way or bad way. Since the column is not a discrete variable anymore, we need use the plot other than box plot. In this case, the scatter plot works perfect for us.

```{r}


p13 <- ggplot(data=final_dataset, 
             aes(x= ST258Q01JA, 
                 y=PV1READ)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(1,5),
                  ylim=c(0,850)) +
    ggtitle("Number of Missing meal vs. Read Performance")+
  theme(plot.title = element_text(size = 6))

p14 <- ggplot(data=final_dataset, 
             aes(x= ST258Q01JA, 
                 y=PV1SCIE)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(1,5),
                  ylim=c(0,850)) +
      ggtitle("Number of Missing meal vs. Science Performance")+
  theme(plot.title = element_text(size = 6))

p15 <- ggplot(data=final_dataset, 
             aes(x= ST258Q01JA, 
                 y=PV1MATH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(1,5),
                  ylim=c(0,850)) +
      ggtitle("Number of Missing meal vs. Math Performance")+
  theme(plot.title = element_text(size = 6))

starv_impact<-p13+p14+p15
starv_impact

```

As the plot shown, the student who do not worry about "having food or not" tend to have higher performance. So we can conclude that, having survival issue have impact on students' performance.

# 3.Conclusion

After the visual analytic about the data set. We can have some basic conclusion on the performance of Singapore student.

-   Over 50% of student can get the grade above average, which means they are doing well

-   Students with family that have enough financial ability that provides stable survival needs and personal room tend to have better performance

-   The performance difference among school are extremely huge, more detailed research are required to find out the reason causing this phenomena
